Handwritten Digit Recognition Using Deep Convolutional Neural Network*

### 2022-04-28 paper/paper.tex
finished the paper

\section{Introduction}
\footnotetext[1]{Code can be found at \href{https://github.com/ammar-elsabe/handwritten-digit-recognition-DCNN.git}{\texttt{github.com/ammar-elsabe/handwritten-\\digit-recognition-DCNN.git}}}
Computer vision is an interdisciplinary field that asks the question: ``Can a computer learn from pictures?''.\@ After all, a picture is worth a thousand words.\@ One interesting field of computer vision is Optical Character Recognition (OCR), which aims to identify the characters in documents. Given the ubiquity of handwritten documents in human transactions, OCR of documents have invaluable practical worth. Optical character recognition is a science that enables to translate various types of documents or images into analyzable, editable and searchable data.
During last decade, researchers have used artificial intelligence / machine learning tools to automatically analyze handwritten and printed documents in order to convert them into electronic format.
An OCR system depends mainly, on the extraction of
features and discrimination/classification of these features
(based on patterns). Handwritten OCR have received increasing attention as a subfield of OCR.\@ It is further categorized into offline\cite{article} and online systems\cite{Connell2001}.
\blindtext

@@

\section{Convolutional Neural Networks}
Convolutional Neural Networks (CNNs) are compriesd of an input layer, hidden layers and an output layer. As with any feed-forward neural network, the middle layers are reffered to as hidden layers because their inputs and outputs are masked by the activation function and final convolution. In a CNN, these hidden layers include layers that perform convolutions, which usually involves computing a dot product of the convolution kernel and the layer's input matrix. The convolution operation generates a feature map as the convolution kernel slides along the input matrix for the layer, which then contributes to the input of the next layer. Other layers such as pooling layers, fully connected layers, and normalization layers are then added.
\subsection{Convolutional Layers}
The input to a CNN is a tensor with the following shape: (number of inputs) x (input height) x (input width) x (input channels). After passing through a convolutional layer, the image is abstracted to a feature map, also known as an activation map, with the following shape: (number of inputs) x (feature map height) x (feature map width) x (feature map channels).
\par Convolutional layers convolve the input and pass the result to the following layer. This is analogous to a neuron's response to a specific stimulus in the visual cortex.

@@

\section{Training the model}

@@

\section{Model Evaluation}

@@

        |C_{x0} \cap C_{y0}| &  \cdots & |C_{x0} \cap C_{y9}| \\
        %|C_{x1} \cap C_{y0}| &  \cdots & |C_{x1} \cap C_{y9}| \\
        \vdots & \ddots & \vdots \\
        |C_{x9} \cap C_{y0}| &  \cdots & |C_{x9} \cap C_{y9}|

@@

In conclusion, we believe that DCNNs will have an important role in upcoming OCR technologies.\@ The results of our experiment are phenomonal, near--human accurate, and with very little cost.\@ There is the issue of our model overfitting too quickly (after just 3 epochs of training), but that can be attributed to the neural network architecture, and with some fine tuning can and has been brought to even more accurate levels.\@

>>>>

\section{Introduction}
\footnotetext[1]{Code can be found at \href{https://github.com/ammar-elsabe/handwritten-digit-recognition-DCNN.git}{\texttt{github.com/ammar-elsabe/handwritten-\\digit-recognition-DCNN.git}}}
Deep learning can now be applied to a growing variety of industries. Convolutional Neural Networking (CNN) is a deep learning technique for evaluating visual data. Convolutional Neural Networking can be used for object detection, face recognition, robotics, sound detection and recognition, video analysis, clustering, pattern recognition, language processing, weather prediction, spam detection, topic categorization, regression analysis, speech recognition, image classification, and handwritten digit recognition.
 Object Character Recognition (OCR) is one of the influencing projects that has used deep learning (Object Character Recognition). OCR is a method of converting printed or written characters into encoded text. The OCR tool scans a document and extracts the information, which is then saved as a digital document. OCR may be implemented in two ways: one by identifying patterns in the characters, and the other by segmenting the text. In any case, machine learning is oblivious to this issue. Handwritten digit recognition (HDR) is a subset of OCR that recognizes digits rather than the entire character's data. HDR is a lighter and quicker alternative to OCR.\@ HDR has a lot of flexibility in areas like medicine, finance, student administration, and the taxes procedure.
Handwritten digit recognition (HDR) refers to a computer's ability to recognize human handwritten digits from a variety of sources, such as photographs, papers, and touch displays, and classify them into ten specified categories (0--9). In the area of deep learning, this has been a topic of inexhaustible investigation. Numerous applications exist for digit recognition, including number plate identification, postal mail sorting, and bank check processing. Because handwritten digit identification is not an optical character recognition, we encounter several obstacles due to the various styles of writing used by different people. A human brain can simply analyse and extract data from a complex image, but for a computer, an image is just a collection of pixels, which are nothing more than a list of values ranging from 0 to 255, with blends such as RGB, B.W., greyscale, etc. Pixel values can be used to extract information from a picture.
This paper experiments with a simple Deep Convolutional Neural Network on the MNIST\cite{6296535} dataset to evaluate their performance.\@ We delve into Convolutional Neural Networks in detail in Section \ref{sec:cnn} as a preliminary to describing the simple architecture used in Section \ref{sec:train}, and we evaluate the the model's performance in Section \ref{sec:test}.

@@

\section{Convolutional Neural Networks}\label{sec:cnn}
Convolutional Neural Networks (CNNs) are compriesd of an input layer, hidden layers and an output layer. As with any feed-forward neural network, the middle layers are reffered to as hidden layers because their inputs and outputs are masked by the activation function and final convolution. In a CNN, these hidden layers include layers that perform convolutions, which usually involves computing a dot product of the convolution kernel and the layer's input matrix. The convolution operation generates a feature map as the convolution kernel slides along the input matrix for the layer, which then contributes to the input of the next layer. Other layers such as pooling layers, fully connected layers, and normalization layers are then added.
\subsection{Convolutional Layers}
The input to a CNN is a tensor with the following shape: (number of inputs) x (input height) x (input width) x (input channels). After passing through a convolutional layer, the image is abstracted to a feature map, also known as an activation map, with the following shape: (number of inputs) x (feature map height) x (feature map width) x (feature map channels).
\par Convolutional layers convolve the input and pass the result to the following layer. This is analogous to a neuron's response to a specific stimulus in the visual cortex.

@@

\section{Training the model}\label{sec:train}

@@

\section{Model Evaluation}\label{sec:test}

@@

        |C_{x0} \cap C_{y0}| & |C_{x0} \cap C_{y1}| & \cdots & |C_{x0} \cap C_{y9}| \\
        |C_{x1} \cap C_{y0}| & |C_{x1} \cap C_{y1}| & \cdots & |C_{x1} \cap C_{y9}| \\
        \vdots & \vdots & \ddots & \vdots \\
        |C_{x9} \cap C_{y0}| & |C_{x9} \cap C_{y1}| & \cdots & |C_{x9} \cap C_{y9}|

@@

In conclusion, we believe that DCNNs have an important role in upcoming OCR technologies.\@ The results of our experiment are phenomonal, near--human accurate, and with very little cost in terms of time spent on developing the model, training the model, and the price of the hardware used.\@
